---
title: "Chart Review Table"
subtitle: "USC Biomedical Student Research Seminar"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: false
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{r packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

library(readxl)
library(dplyr)
library(purrr)
library(tibble)
library(gtsummary)
library(gt)
library(ggplot2)

```

# Introduction

This code was generated for the **2026 Biomedical Research Seminar** at the University of Southern California by Jonathan Nelson. It creates a descriptive summary table comparing patient, perioperative, pathologic, and postoperative outcomes between surgical approaches, and includes p-values and statistical test labels for each comparison.

The analysis uses a de-identified dataset provided by Sammie Rosen. The dataset includes demographic, clinical, operative, histologic, and follow-up variables collected for patients undergoing **extraperitoneal versus transvesical simple prostatectomy**.

## Download folder with files for analysis

[**Download the Folder**](data/chart_review.zip)

## Load in the Dataset

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- read_excel("MASTER EP vs TV simples - deidentified.xlsx", sheet = "Sheet1")

df

```

## Name and Factor Table

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- read_excel("MASTER EP vs TV simples - deidentified.xlsx", sheet = "Sheet1") %>%
  mutate(
    approach = factor(
      `Surgical Approach (1=EP or 2=TV)`,
      levels = c(1, 2),
      labels = c("Extraperitoneal", "Transvesical")
    ),

    `HTN status (0=No or 1=Yes)` = factor(`HTN status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `DM2 status (0=No or 1=Yes)` = factor(`DM2 status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `HLD (0=No or 1=Yes)` = factor(`HLD (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)` = factor(`prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Prior Abdominal Surgery (0=No or 1=Yes)` = factor(`Prior Abdominal Surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Pre-Op Biopsy Done (0=No 1=Yes)` = factor(`Pre-Op Biopsy Done (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Pre-Op Biopsy Result (0=Neg 1=Pos)` = factor(`Pre-Op Biopsy Result (0=Neg 1=Pos)`, levels = c(0, 1), labels = c("Negative", "Positive")),
    `Pre-Op Foley Present (0=No 1=Yes)` = factor(`Pre-Op Foley Present (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Lymph Node Dissection (0=No or 1=Yes)` = factor(`Lymph Node Dissection (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Conversion (0=No or 1=Yes)` = factor(`Conversion (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Drain placement (0=No or 1=Yes)` = factor(`Drain placement (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `pRBCs given (0=No 1=Yes)` = factor(`pRBCs given (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `LOS (0=Same Day 1=Inpatient)` = factor(`LOS (0=Same Day 1=Inpatient)`, levels = c(0, 1), labels = c("Same Day", "Inpatient")),
    `Narcotics (PO day 0-1) (0 = no, 1 = yes)` = factor(`Narcotics (PO day 0-1) (0 = no, 1 = yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30-Day Readmission (0=No or 1=Yes)` = factor(`30-Day Readmission (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad` =
      factor(`Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad`,
             levels = c(0, 1), labels = c("No", "Yes")),
    `Continence 3 Months (0=No or 1=Yes)` = factor(`Continence 3 Months (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30 day Complications (0=no 1=Yes)` = factor(`30 day Complications (0=no 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `ASA (1-5)` = factor(`ASA (1-5)`)
  )

df 

```

## Let's Look at One of the Variables

### Age at Time of Surgery

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df$`Age (at time of surgery)`

```

#### Calculate the Statistic

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df %>%
  group_by(approach) %>%
  summarise(
    mean_age = mean(`Age (at time of surgery)`, na.rm = TRUE),
    sd_age   = sd(`Age (at time of surgery)`, na.rm = TRUE),
    .groups = "drop"
  )

wilcox.test(`Age (at time of surgery)` ~ approach, data = df)

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

summary_df <- df %>%
  group_by(approach) %>%
  summarise(
    mean_age = mean(`Age (at time of surgery)`, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(df, aes(x = approach, y = `Age (at time of surgery)`)) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 4) +
  geom_point(
    data = summary_df,
    aes(x = approach, y = mean_age),
    color = "red",
    size = 6
  ) +
  labs(
    x = "Approach",
    y = "Age (at time of surgery)",
    title = "Age Distribution by Surgical Approach (with Means)"
  ) +
  theme_classic()

```

#### Putting It Into a Table

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Summary stats
age_summary <- df %>%
  group_by(approach) %>%
  summarise(
    mean = mean(`Age (at time of surgery)`, na.rm = TRUE),
    sd   = sd(`Age (at time of surgery)`, na.rm = TRUE),
    .groups = "drop"
  )

# Format into one row
age_tbl <- tibble(
  Variable = "Age (yr)",
  Extraperitoneal = sprintf("%.1f (%.1f)",
                            age_summary$mean[age_summary$approach == "Extraperitoneal"],
                            age_summary$sd[age_summary$approach == "Extraperitoneal"]),
  Transvesical = sprintf("%.1f (%.1f)",
                         age_summary$mean[age_summary$approach == "Transvesical"],
                         age_summary$sd[age_summary$approach == "Transvesical"]),
  `P-value` = formatC(
    wilcox.test(`Age (at time of surgery)` ~ approach, data = df)$p.value,
    format = "f", digits = 3
  ),
  Test = "Wilcoxon rank-sum"
)

age_tbl %>%
  gt()

```

## Working Table

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- read_excel("MASTER EP vs TV simples - deidentified.xlsx", sheet = "Sheet1") %>%
  mutate(
    approach = factor(
      `Surgical Approach (1=EP or 2=TV)`,
      levels = c(1, 2),
      labels = c("Extraperitoneal", "Transvesical")
    ),

    `HTN status (0=No or 1=Yes)` = factor(`HTN status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `DM2 status (0=No or 1=Yes)` = factor(`DM2 status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `HLD (0=No or 1=Yes)` = factor(`HLD (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)` = factor(`prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Prior Abdominal Surgery (0=No or 1=Yes)` = factor(`Prior Abdominal Surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Pre-Op Biopsy Done (0=No 1=Yes)` = factor(`Pre-Op Biopsy Done (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Pre-Op Biopsy Result (0=Neg 1=Pos)` = factor(`Pre-Op Biopsy Result (0=Neg 1=Pos)`, levels = c(0, 1), labels = c("Negative", "Positive")),
    `Pre-Op Foley Present (0=No 1=Yes)` = factor(`Pre-Op Foley Present (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Lymph Node Dissection (0=No or 1=Yes)` = factor(`Lymph Node Dissection (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Conversion (0=No or 1=Yes)` = factor(`Conversion (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Drain placement (0=No or 1=Yes)` = factor(`Drain placement (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `pRBCs given (0=No 1=Yes)` = factor(`pRBCs given (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `LOS (0=Same Day 1=Inpatient)` = factor(`LOS (0=Same Day 1=Inpatient)`, levels = c(0, 1), labels = c("Same Day", "Inpatient")),
    `Narcotics (PO day 0-1) (0 = no, 1 = yes)` = factor(`Narcotics (PO day 0-1) (0 = no, 1 = yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30-Day Readmission (0=No or 1=Yes)` = factor(`30-Day Readmission (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad` =
      factor(`Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad`,
             levels = c(0, 1), labels = c("No", "Yes")),
    `Continence 3 Months (0=No or 1=Yes)` = factor(`Continence 3 Months (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30 day Complications (0=no 1=Yes)` = factor(`30 day Complications (0=no 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `ASA (1-5)` = factor(`ASA (1-5)`)
  )

baseline_labels <- list(
  `Age (at time of surgery)` ~ "Age (yr)",
  BMI ~ "BMI (kg/m^2)",
  CCI ~ "CCI",
  `ASA (1-5)` ~ "ASA physical status grading",
  `HTN status (0=No or 1=Yes)` ~ "History of hypertension",
  `DM2 status (0=No or 1=Yes)` ~ "History of type 2 diabetes mellitus",
  `HLD (0=No or 1=Yes)` ~ "History of hyperlipidemia",
  `prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)` ~ "History of prior inguinal/pelvic surgery",
  `Prior Abdominal Surgery (0=No or 1=Yes)` ~ "History of prior abdominal surgery",
  `Pre-op Hgb` ~ "Pre-op Hgb",
  `Pre-op PSA` ~ "Pre-op PSA (ng/mL)",
  `Pre-op Prostate Volume (g)` ~ "Prostate volume (ml)"
)

preop_labels <- list(
  `SHIM Score (pre-op)` ~ "SHIM",
  `IPSS (pre-op)` ~ "IPSS",
  `QOL (pre-op)` ~ "QoL",
  `Pre-Op Biopsy Done (0=No 1=Yes)` ~ "Biopsy done",
  `Pre-Op Biopsy Result (0=Neg 1=Pos)` ~ "Malignancy on biopsy",
  `Pre-Op Foley Present (0=No 1=Yes)` ~ "Pre-op foley placed"
)

intraop_labels <- list(
  `Console Time (minutes)` ~ "Console time (minutes)",
  `Operative Time (minutes)` ~ "Operative time (minutes)",
  `EBL (cc)` ~ "Estimated blood loss (mL)",
  `Conversion (0=No or 1=Yes)` ~ "Conversion to multiport",
  `Lymph Node Dissection (0=No or 1=Yes)` ~ "Lymph node dissection",
  `Drain placement (0=No or 1=Yes)` ~ "Drain placement",
  `pRBCs given (0=No 1=Yes)` ~ "Blood units given",
  `delta Hgb` ~ "dHgb (%)"
)

histology_labels <- list(
  `Specimen Weight (g)` ~ "Specimen weight (g)",
  `Specimen weight/pre-op estmated weight` ~ "Specimen wt/pre-op estimated wt",
  `Pathologic Diagnosis (incidental malignancy) 0= benign 1= malignant` ~ "Incidental malignancy finding"
)

postop_care_labels <- list(
  `LOS (0=Same Day 1=Inpatient)` ~ "Length of stay <24 hours",
  `LOS (hours) (look at note times)` ~ "Average length of stay (hours)",
  `Narcotics (PO day 0-1) (0 = no, 1 = yes)` ~ "Narcotics use post-op day 0-1",
  `Postop Pain Score (PACU)` ~ "Post Op Pain Score",
  `Foley Duration (Days post-op)` ~ "Foley catheter duration (days)"
)

outcome_labels <- list(
  `30 day Complications (0=no 1=Yes)` ~ "30 day postoperative complications (Clavien >=2 only)",
  `30-Day Readmission (0=No or 1=Yes)` ~ "30 day readmission",
  `Post-op PVR` ~ "Post-op PVR",
  `Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad` ~ "Early Continence at 4 weeks",
  `Continence 3 Months (0=No or 1=Yes)` ~ "Continence at 3 months"
)

format_p <- function(x) {
  dplyr::case_when(
    is.na(x) ~ NA_character_,
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(x, format = "f", digits = 3)
  )
}

safe_pvalue <- function(data, group_var, var, continuous = FALSE) {
  x <- data[[var]]
  g <- data[[group_var]]
  keep <- complete.cases(x, g)
  x <- x[keep]
  g <- g[keep]

  if (length(unique(g)) < 2) return(NA_real_)

  tryCatch({
    if (continuous) {
      wilcox.test(x ~ g)$p.value
    } else {
      tab <- table(x, g)
      if (any(dim(tab) < 2)) return(NA_real_)
      if (all(dim(tab) == c(2, 2))) {
        fisher.test(tab)$p.value
      } else {
        suppressWarnings(chisq.test(tab, correct = FALSE)$p.value)
      }
    }
  }, error = function(e) NA_real_)
}

calc_manual_pvals <- function(data, group_var, vars, continuous_vars) {
  map_dfr(vars, function(v) {
    tibble(
      variable = v,
      manual_p = safe_pvalue(data, group_var, v, continuous = v %in% continuous_vars)
    )
  })
}

attach_manual_pvals <- function(tbl, pvals) {
  tbl %>%
    modify_table_body(
      ~ .x %>%
        left_join(pvals, by = "variable") %>%
        mutate(
          p.value = if_else(
            row_type == "label",
            format_p(manual_p),
            NA_character_
          )
        ) %>%
        select(-manual_p)
    ) %>%
    modify_header(p.value ~ "**P-value**") %>%
    modify_column_unhide(p.value)
}

make_tbl <- function(data, vars, labels, statistic, continuous_vars) {
  pvals <- calc_manual_pvals(data, "approach", vars, continuous_vars)

  data %>%
    select(approach, all_of(vars)) %>%
    tbl_summary(
      by = approach,
      label = labels,
      statistic = statistic,
      digits = all_continuous() ~ 1,
      missing = "ifany"
    ) %>%
    attach_manual_pvals(pvals)
}

baseline_vars <- c(
  "Age (at time of surgery)", "BMI", "CCI", "ASA (1-5)",
  "HTN status (0=No or 1=Yes)", "DM2 status (0=No or 1=Yes)", "HLD (0=No or 1=Yes)",
  "prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)",
  "Prior Abdominal Surgery (0=No or 1=Yes)",
  "Pre-op Hgb", "Pre-op PSA", "Pre-op Prostate Volume (g)"
)
baseline_cont <- c(
  "Age (at time of surgery)", "BMI", "CCI",
  "Pre-op Hgb", "Pre-op PSA", "Pre-op Prostate Volume (g)"
)

preop_vars <- c(
  "SHIM Score (pre-op)", "IPSS (pre-op)", "QOL (pre-op)",
  "Pre-Op Biopsy Done (0=No 1=Yes)",
  "Pre-Op Biopsy Result (0=Neg 1=Pos)",
  "Pre-Op Foley Present (0=No 1=Yes)"
)
preop_cont <- c("SHIM Score (pre-op)", "IPSS (pre-op)", "QOL (pre-op)")

intraop_vars <- c(
  "Console Time (minutes)", "Operative Time (minutes)", "EBL (cc)",
  "Conversion (0=No or 1=Yes)",
  "Lymph Node Dissection (0=No or 1=Yes)",
  "Drain placement (0=No or 1=Yes)",
  "pRBCs given (0=No 1=Yes)",
  "delta Hgb"
)
intraop_cont <- c("Console Time (minutes)", "Operative Time (minutes)", "EBL (cc)", "delta Hgb")

histology_vars <- c(
  "Specimen Weight (g)",
  "Specimen weight/pre-op estmated weight",
  "Pathologic Diagnosis (incidental malignancy) 0= benign 1= malignant"
)
histology_cont <- c("Specimen Weight (g)", "Specimen weight/pre-op estmated weight")

postop_care_vars <- c(
  "LOS (0=Same Day 1=Inpatient)",
  "LOS (hours) (look at note times)",
  "Narcotics (PO day 0-1) (0 = no, 1 = yes)",
  "Postop Pain Score (PACU)",
  "Foley Duration (Days post-op)"
)
postop_care_cont <- c("LOS (hours) (look at note times)", "Postop Pain Score (PACU)", "Foley Duration (Days post-op)")

outcome_vars <- c(
  "30 day Complications (0=no 1=Yes)",
  "30-Day Readmission (0=No or 1=Yes)",
  "Post-op PVR",
  "Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad",
  "Continence 3 Months (0=No or 1=Yes)"
)
outcome_cont <- c("Post-op PVR")

baseline_tbl <- make_tbl(
  df,
  vars = baseline_vars,
  labels = baseline_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    `ASA (1-5)` ~ "{n} ({p}%)",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = baseline_cont
)

preop_tbl <- make_tbl(
  df,
  vars = preop_vars,
  labels = preop_labels,
  statistic = list(
    c(`SHIM Score (pre-op)`, `IPSS (pre-op)`, `QOL (pre-op)`) ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = preop_cont
)

intraop_tbl <- make_tbl(
  df,
  vars = intraop_vars,
  labels = intraop_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = intraop_cont
)

histology_tbl <- make_tbl(
  df,
  vars = histology_vars,
  labels = histology_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = histology_cont
)

postop_care_tbl <- make_tbl(
  df,
  vars = postop_care_vars,
  labels = postop_care_labels,
  statistic = list(
    `LOS (hours) (look at note times)` ~ "{mean} ({sd})",
    `Postop Pain Score (PACU)` ~ "{mean} ({sd})",
    `Foley Duration (Days post-op)` ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = postop_care_cont
)

outcomes_tbl <- make_tbl(
  df,
  vars = outcome_vars,
  labels = outcome_labels,
  statistic = list(
    `Post-op PVR` ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = outcome_cont
)

final_tbl <- tbl_stack(
  tbls = list(
    baseline_tbl,
    preop_tbl,
    intraop_tbl,
    histology_tbl,
    postop_care_tbl,
    outcomes_tbl
  ),
  group_header = c(
    "Baseline characteristics",
    "Pre-operative variables",
    "Intraoperative parameters",
    "Histopathology",
    "Postoperative care",
    "Postoperative outcomes"
  )
) %>%
  modify_column_unhide(p.value) %>%
  bold_labels()

final_tbl


```

## Breaking Down the Code

This script: <br>
     1. Reads the data from Excel  <br>
     2. Recodes raw 0/1 variables into readable categories  <br>
     3. Groups variables into clinical sections  <br>
     4. Builds summary statistics for each section  <br>
     5. Computes p-values comparing the two surgical approaches  <br>
     6. Combines everything into one polished table  <br>

### Read the Excel File and Create New Variable

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- read_excel(...) %>%
  mutate(...)

```

### Create Label Lists

baseline_labels <br>
preop_labels <br>
intraop_labels <br>
histology_labels <br>
postop_care_labels <br>
outcome_labels <br>

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

`Age (at time of surgery)` ~ "Age (yr)"

```

### Format the p-value

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

format_p <- function(x) { ... }

```

### Compute p-value safely

This is the section that was the most challenging.

#### Wilcox for continuous variable
#### Fisher/Chi-squared for categorical variable

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

safe_pvalue <- function(data, group_var, var, continuous = FALSE) { ... }

wilcox.test(x ~ g)$p.value

tab <- table(x, g)

```

### Build a p-value table for many variables

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

calc_manual_pvals <- function(data, group_var, vars, continuous_vars) { ... }

```

### Attach p-values into the gtsummary table

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

attach_manual_pvals <- function(tbl, pvals) { ... }

```

### Create a reusable table-making function

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

make_tbl <- function(data, vars, labels, statistic, continuous_vars) { ... }

```

### Define which variables belong in each section

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

make_tbl <- function(data, vars, labels, statistic, continuous_vars) { ... }

```

### Create each summary table section

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

baseline_tbl <- make_tbl(...)

```

### Stack all sections into one final table

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

final_tbl <- tbl_stack(
  tbls = list(...)
  group_header = c(...)
)

```

### Display the final table

```{r echo=TRUE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

final_tbl)

```

## Table with Statistics Info

I wanted an indication of which statistic test was used for each comparison.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- read_excel("MASTER EP vs TV simples - deidentified.xlsx", sheet = "Sheet1") %>%
  mutate(
    approach = factor(
      `Surgical Approach (1=EP or 2=TV)`,
      levels = c(1, 2),
      labels = c("Extraperitoneal", "Transvesical")
    ),

    `HTN status (0=No or 1=Yes)` = factor(`HTN status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `DM2 status (0=No or 1=Yes)` = factor(`DM2 status (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `HLD (0=No or 1=Yes)` = factor(`HLD (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)` = factor(`prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Prior Abdominal Surgery (0=No or 1=Yes)` = factor(`Prior Abdominal Surgery (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Pre-Op Biopsy Done (0=No 1=Yes)` = factor(`Pre-Op Biopsy Done (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Pre-Op Biopsy Result (0=Neg 1=Pos)` = factor(`Pre-Op Biopsy Result (0=Neg 1=Pos)`, levels = c(0, 1), labels = c("Negative", "Positive")),
    `Pre-Op Foley Present (0=No 1=Yes)` = factor(`Pre-Op Foley Present (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `Lymph Node Dissection (0=No or 1=Yes)` = factor(`Lymph Node Dissection (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Conversion (0=No or 1=Yes)` = factor(`Conversion (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Drain placement (0=No or 1=Yes)` = factor(`Drain placement (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `pRBCs given (0=No 1=Yes)` = factor(`pRBCs given (0=No 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `LOS (0=Same Day 1=Inpatient)` = factor(`LOS (0=Same Day 1=Inpatient)`, levels = c(0, 1), labels = c("Same Day", "Inpatient")),
    `Narcotics (PO day 0-1) (0 = no, 1 = yes)` = factor(`Narcotics (PO day 0-1) (0 = no, 1 = yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30-Day Readmission (0=No or 1=Yes)` = factor(`30-Day Readmission (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad` =
      factor(`Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad`,
             levels = c(0, 1), labels = c("No", "Yes")),
    `Continence 3 Months (0=No or 1=Yes)` = factor(`Continence 3 Months (0=No or 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),
    `30 day Complications (0=no 1=Yes)` = factor(`30 day Complications (0=no 1=Yes)`, levels = c(0, 1), labels = c("No", "Yes")),

    `ASA (1-5)` = factor(`ASA (1-5)`)
  )

baseline_labels <- list(
  `Age (at time of surgery)` ~ "Age (yr)",
  BMI ~ "BMI (kg/m^2)",
  CCI ~ "CCI",
  `ASA (1-5)` ~ "ASA physical status grading",
  `HTN status (0=No or 1=Yes)` ~ "History of hypertension",
  `DM2 status (0=No or 1=Yes)` ~ "History of type 2 diabetes mellitus",
  `HLD (0=No or 1=Yes)` ~ "History of hyperlipidemia",
  `prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)` ~ "History of prior inguinal/pelvic surgery",
  `Prior Abdominal Surgery (0=No or 1=Yes)` ~ "History of prior abdominal surgery",
  `Pre-op Hgb` ~ "Pre-op Hgb",
  `Pre-op PSA` ~ "Pre-op PSA (ng/mL)",
  `Pre-op Prostate Volume (g)` ~ "Prostate volume (ml)"
)

preop_labels <- list(
  `SHIM Score (pre-op)` ~ "SHIM",
  `IPSS (pre-op)` ~ "IPSS",
  `QOL (pre-op)` ~ "QoL",
  `Pre-Op Biopsy Done (0=No 1=Yes)` ~ "Biopsy done",
  `Pre-Op Biopsy Result (0=Neg 1=Pos)` ~ "Malignancy on biopsy",
  `Pre-Op Foley Present (0=No 1=Yes)` ~ "Pre-op foley placed"
)

intraop_labels <- list(
  `Console Time (minutes)` ~ "Console time (minutes)",
  `Operative Time (minutes)` ~ "Operative time (minutes)",
  `EBL (cc)` ~ "Estimated blood loss (mL)",
  `Conversion (0=No or 1=Yes)` ~ "Conversion to multiport",
  `Lymph Node Dissection (0=No or 1=Yes)` ~ "Lymph node dissection",
  `Drain placement (0=No or 1=Yes)` ~ "Drain placement",
  `pRBCs given (0=No 1=Yes)` ~ "Blood units given",
  `delta Hgb` ~ "dHgb (%)"
)

histology_labels <- list(
  `Specimen Weight (g)` ~ "Specimen weight (g)",
  `Specimen weight/pre-op estmated weight` ~ "Specimen wt/pre-op estimated wt",
  `Pathologic Diagnosis (incidental malignancy) 0= benign 1= malignant` ~ "Incidental malignancy finding"
)

postop_care_labels <- list(
  `LOS (0=Same Day 1=Inpatient)` ~ "Length of stay <24 hours",
  `LOS (hours) (look at note times)` ~ "Average length of stay (hours)",
  `Narcotics (PO day 0-1) (0 = no, 1 = yes)` ~ "Narcotics use post-op day 0-1",
  `Postop Pain Score (PACU)` ~ "Post Op Pain Score",
  `Foley Duration (Days post-op)` ~ "Foley catheter duration (days)"
)

outcome_labels <- list(
  `30 day Complications (0=no 1=Yes)` ~ "30 day postoperative complications (Clavien >=2 only)",
  `30-Day Readmission (0=No or 1=Yes)` ~ "30 day readmission",
  `Post-op PVR` ~ "Post-op PVR",
  `Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad` ~ "Early Continence at 4 weeks",
  `Continence 3 Months (0=No or 1=Yes)` ~ "Continence at 3 months"
)

format_p <- function(x) {
  dplyr::case_when(
    is.na(x) ~ NA_character_,
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(x, format = "f", digits = 3)
  )
}

safe_test <- function(data, group_var, var, continuous = FALSE) {
  x <- data[[var]]
  g <- data[[group_var]]
  keep <- complete.cases(x, g)
  x <- x[keep]
  g <- g[keep]

  if (length(unique(g)) < 2) {
    return(list(p = NA_real_, test = NA_character_))
  }

  tryCatch({
    if (continuous) {
      p <- wilcox.test(x ~ g)$p.value
      list(p = p, test = "Wilcoxon rank-sum")
    } else {
      tab <- table(x, g)
      if (any(dim(tab) < 2)) {
        return(list(p = NA_real_, test = NA_character_))
      }
      if (all(dim(tab) == c(2, 2))) {
        p <- fisher.test(tab)$p.value
        list(p = p, test = "Fisher's exact")
      } else {
        p <- suppressWarnings(chisq.test(tab, correct = FALSE)$p.value)
        list(p = p, test = "Chi-square")
      }
    }
  }, error = function(e) {
    list(p = NA_real_, test = NA_character_)
  })
}

calc_manual_stats <- function(data, group_var, vars, continuous_vars) {
  map_dfr(vars, function(v) {
    res <- safe_test(data, group_var, v, continuous = v %in% continuous_vars)
    tibble(
      variable = v,
      manual_p = res$p,
      manual_test = res$test
    )
  })
}

attach_manual_stats <- function(tbl, stats) {
  tbl %>%
    modify_table_body(
      ~ .x %>%
        left_join(stats, by = "variable") %>%
        mutate(
          p.value = if_else(row_type == "label", format_p(manual_p), NA_character_),
          stat_test = if_else(row_type == "label", manual_test, NA_character_)
        ) %>%
        select(-manual_p, -manual_test)
    ) %>%
    modify_header(
      p.value ~ "**P-value**",
      stat_test ~ "**Test used**"
    ) %>%
    modify_column_unhide(c(p.value, stat_test))
}

make_tbl <- function(data, vars, labels, statistic, continuous_vars) {
  stats <- calc_manual_stats(data, "approach", vars, continuous_vars)

  data %>%
    select(approach, all_of(vars)) %>%
    tbl_summary(
      by = approach,
      label = labels,
      statistic = statistic,
      digits = all_continuous() ~ 1,
      missing = "ifany"
    ) %>%
    attach_manual_stats(stats)
}

baseline_vars <- c(
  "Age (at time of surgery)", "BMI", "CCI", "ASA (1-5)",
  "HTN status (0=No or 1=Yes)", "DM2 status (0=No or 1=Yes)", "HLD (0=No or 1=Yes)",
  "prior Inguinal Hernia/pelvic surgery (0=No or 1=Yes)",
  "Prior Abdominal Surgery (0=No or 1=Yes)",
  "Pre-op Hgb", "Pre-op PSA", "Pre-op Prostate Volume (g)"
)
baseline_cont <- c(
  "Age (at time of surgery)", "BMI", "CCI",
  "Pre-op Hgb", "Pre-op PSA", "Pre-op Prostate Volume (g)"
)

preop_vars <- c(
  "SHIM Score (pre-op)", "IPSS (pre-op)", "QOL (pre-op)",
  "Pre-Op Biopsy Done (0=No 1=Yes)",
  "Pre-Op Biopsy Result (0=Neg 1=Pos)",
  "Pre-Op Foley Present (0=No 1=Yes)"
)
preop_cont <- c("SHIM Score (pre-op)", "IPSS (pre-op)", "QOL (pre-op)")

intraop_vars <- c(
  "Console Time (minutes)", "Operative Time (minutes)", "EBL (cc)",
  "Conversion (0=No or 1=Yes)",
  "Lymph Node Dissection (0=No or 1=Yes)",
  "Drain placement (0=No or 1=Yes)",
  "pRBCs given (0=No 1=Yes)",
  "delta Hgb"
)
intraop_cont <- c("Console Time (minutes)", "Operative Time (minutes)", "EBL (cc)", "delta Hgb")

histology_vars <- c(
  "Specimen Weight (g)",
  "Specimen weight/pre-op estmated weight",
  "Pathologic Diagnosis (incidental malignancy) 0= benign 1= malignant"
)
histology_cont <- c("Specimen Weight (g)", "Specimen weight/pre-op estmated weight")

postop_care_vars <- c(
  "LOS (0=Same Day 1=Inpatient)",
  "LOS (hours) (look at note times)",
  "Narcotics (PO day 0-1) (0 = no, 1 = yes)",
  "Postop Pain Score (PACU)",
  "Foley Duration (Days post-op)"
)
postop_care_cont <- c("LOS (hours) (look at note times)", "Postop Pain Score (PACU)", "Foley Duration (Days post-op)")

outcome_vars <- c(
  "30 day Complications (0=no 1=Yes)",
  "30-Day Readmission (0=No or 1=Yes)",
  "Post-op PVR",
  "Early Continence at 6 Weeks (0=No or 1=Yes) continent = no pads or 1 protection pad",
  "Continence 3 Months (0=No or 1=Yes)"
)
outcome_cont <- c("Post-op PVR")

baseline_tbl <- make_tbl(
  df,
  vars = baseline_vars,
  labels = baseline_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    `ASA (1-5)` ~ "{n} ({p}%)",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = baseline_cont
)

preop_tbl <- make_tbl(
  df,
  vars = preop_vars,
  labels = preop_labels,
  statistic = list(
    c(`SHIM Score (pre-op)`, `IPSS (pre-op)`, `QOL (pre-op)`) ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = preop_cont
)

intraop_tbl <- make_tbl(
  df,
  vars = intraop_vars,
  labels = intraop_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = intraop_cont
)

histology_tbl <- make_tbl(
  df,
  vars = histology_vars,
  labels = histology_labels,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = histology_cont
)

postop_care_tbl <- make_tbl(
  df,
  vars = postop_care_vars,
  labels = postop_care_labels,
  statistic = list(
    `LOS (hours) (look at note times)` ~ "{mean} ({sd})",
    `Postop Pain Score (PACU)` ~ "{mean} ({sd})",
    `Foley Duration (Days post-op)` ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = postop_care_cont
)

outcomes_tbl <- make_tbl(
  df,
  vars = outcome_vars,
  labels = outcome_labels,
  statistic = list(
    `Post-op PVR` ~ "{median} ({p25}, {p75})",
    all_categorical() ~ "{n} ({p}%)"
  ),
  continuous_vars = outcome_cont
)

final_tbl_2 <- tbl_stack(
  tbls = list(
    baseline_tbl,
    preop_tbl,
    intraop_tbl,
    histology_tbl,
    postop_care_tbl,
    outcomes_tbl
  ),
  group_header = c(
    "Baseline characteristics",
    "Pre-operative variables",
    "Intraoperative parameters",
    "Histopathology",
    "Postoperative care",
    "Postoperative outcomes"
  )
) %>%
  modify_column_unhide(c(p.value, stat_test)) %>%
  bold_labels()

final_tbl_2

```

## Improving Aesthetics of Table

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


final_tbl %>%
  as_gt() %>%
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold"),
      gt::cell_borders(
        sides = c("top", "bottom"),
        color = "black",
        weight = gt::px(2)
      )
    ),
    locations = gt::cells_row_groups()
  ) %>%
  gt::tab_options(
    row_group.as_column = FALSE
  )

```

## Revise the Table Your Way

1) Make the categories with significant differences red?
2) Bold significant P values?
3) Fix the kg/m^2 to actually be superscript in table?

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}




```

# Session Info

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Sys.time()

sessionInfo()

```

