---
title: "Learning R through IRIS"
subtitle: "USC Biomedical Student Research Seminar"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: false
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



```{r packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

library(dplyr)
library(ggplot2)
library(tibble)
library(viridis)
library(RColorBrewer)

```

# Introduction

This tutorial uses the built-in `iris` dataset to introduce core skills for working with data in R. Rather than focusing on abstract concepts, we will learn by exploring a real dataset step by step.

You will practice using `dplyr` to organize and transform data, helping you focus on the parts of the dataset that matter for your questions. You will then use `ggplot2` to create clear and informative visualizations, building from simple plots to more detailed, multi-dimensional graphics with customized color palettes.

Finally, you will connect these visual patterns to statistical tests, learning how to formally evaluate differences and relationships in the data.

By the end of this tutorial, you should be able to: <br>
- explore and summarize a dataset <br>
- reshape and refine data for analysis using `dplyr` <br>
- create effective visualizations with `ggplot2` <br>
- apply basic statistical tests and interpret their results <br>

The goal is not just to learn individual tools, but to see how they work together as part of a complete data analysis workflow.

## **Download** this .Rmd file

[**Download RMD file**](data/learning_r_through_iris.zip)

## **Read** in the dataset

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

data(iris)

```

## **Inspect** the data

`head` gives the top 6 rows of the dataframe

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

head(iris)

```

## Check the **Structure**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

str(iris)

```

## Get a **Summary**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

summary(iris)

```

# Basic `dplyr` summaries

## Count **number** of entries by **Species**

In English: "take the object iris and then count the groups in the column Species"

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris %>%
  count(Species)

```
### Compute **average measurements** by species

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris %>%
  group_by(Species) %>%
  summarise(
    n = n(),
    avg_sepal_length = mean(Sepal.Length),
    avg_sepal_width  = mean(Sepal.Width),
    avg_petal_length = mean(Petal.Length),
    avg_petal_width  = mean(Petal.Width)
  )

```

## **Filter** rows

### Look only at `setosa`

In English: "take the object iris and then filter to only include Species = `setosa` rows, and then show the top 6 rows"

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris %>%
  filter(Species == "setosa") %>%
  head()

```

### Look only at flowers with **large petals**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris %>%
  filter(Petal.Length > 5, Petal.Width > 1.5) %>%
  head()

```

## Create **new variables**

In English: "Create the new object `iris2` by starting with the object `iris` and then creating a new column with the name `Petal.Area` that is the value of `Petal.Length * Petal.Width`"

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris2 <- iris %>%
  mutate(Petal.Area = Petal.Length * Petal.Width)

head(iris2)

```

### Create and Summarize

How do you say this in english?

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris2 <- iris2 %>%
  mutate(Sepal.Ratio = Sepal.Length / Sepal.Width)

summary(iris2$Sepal.Ratio)

```

## First **ggplot**: a scatterplot

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point()

```

### **Color** by species

`color = Species`//Also adding **custom plot labels**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  labs(
    title = "Sepal Length vs Sepal Width",
    subtitle = "Colored by species",
    x = "Sepal Length",
    y = "Sepal Width"
  )

```

#### Color palette = `RColorBrewer`

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Petal Dimensions by Species",
    x = "Petal Length",
    y = "Petal Width"
  )

```

#### Colorblind-friendly = `viridis`

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = "Petal Dimensions by Species",
    x = "Petal Length",
    y = "Petal Width"
  )
```


## Multidimensional Plots

### **Going 4D!**

x = Sepal.Length //
y = Petal.Length //
color = Species //
size = Petal.Width

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris2, aes(
  x = Sepal.Length,
  y = Petal.Length,
  color = Species,
  size = Petal.Width
)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    title = "A Multidimensional Plot",
    subtitle = "Color = species, size = petal width",
    x = "Sepal Length",
    y = "Petal Length"
  ) +
  theme_minimal()


```

## **Faceting** for comparison

`facet_wrap(~Species)`

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12, fig.height=4}

ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(aes(color = Species), size = 2) +
  facet_wrap(~Species) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = "Petal Size by Species",
    x = "Petal Length",
    y = "Petal Width"
  )

```

## **Summaries** plus plots

### Calculate the Summary

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

species_summary <- iris %>%
  group_by(Species) %>%
  summarise(
    avg_petal_width = mean(Petal.Width),
    avg_petal_length = mean(Petal.Length)
  )

species_summary

```

### Plot the **Summary**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(species_summary, aes(x = Species, y = avg_petal_width, fill = Species)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    title = "Average Petal Width by Species",
    x = "Species",
    y = "Average Petal Width"
  ) +
  theme(legend.position = "none")

```

# Statistics

## t-test

### Filter to **two variables**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

iris_two <- iris %>%
  filter(Species %in% c("setosa", "versicolor"))

t.test(Petal.Length ~ Species, data = iris_two)

```

### **Check the Plot**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris_two, aes(x = Species, y = Petal.Length, color = Species)) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Petal Length by Species")

```

## ANOVA

### For more than two variables

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

anova_model <- aov(Petal.Length ~ Species, data = iris)
summary(anova_model)

```

### Pairwise comparisons (Tukey HSD)

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

TukeyHSD(anova_model)

```

### **Check the Plot**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Species, y = Petal.Length, color = Species)) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Petal Length by Species")

```


## Correlation

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

cor.test(iris$Petal.Length, iris$Petal.Width)

```

### **Check the Plot**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Petal Length vs Petal Width")

```


# Build Your Own Graph


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}



```


# Session Info

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Sys.time()

sessionInfo()

```

